How is active school travel framed in Ontario, Canada?

### 2022-09-07 AST-Framing-Ontario/AST-Framing-Ontario.tex
Edited for submission to TR-A Policy and Practice

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{Social Science \& Medicine} % Sets Journal name

@@

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={How is active school travel framed in Ontario, Canada?},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}

@@

% Pandoc citation processing
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

  \title{How is active school travel framed in Ontario, Canada?}
    \author[Some Department]{Author 1\corref{1}}
   \ead{author1@example.com} 
    \author[Some Department]{Author 2}
   \ead{author2@example.com} 
    \author[Another University]{Author 3\corref{2}}
   \ead{author3@example.com} 
    \author[Some Institute]{Author 4\corref{2}}
   \ead{author4@example.com} 
    \author[Some University]{Author 5}
   \ead{author5@example.com} 
    \author[Some Department]{Author 6}
   \ead{author6@example.com} 
      \address[Some Department]{Department, Street, City, Province,
Postal Code}
    \address[Another University]{Department, Street, City, Province,
Postal Code}
    \address[Some Institute]{Street, City, Province, Postal Code}
    \address[Some University]{Department, Street, City, Province, Postal
Code}
      \cortext[1]{Corresponding Author}
    \cortext[2]{Equal contribution}

  \begin{abstract}
  Active school travel (AST) has been increasingly encouraged by various
  stakeholders in Ontario, Canada through efforts such as school travel
  planning. Education strategies like workshops or resources that
  promote AST are commonly implemented. The framing of AST through such
  strategies may influence how walking and bicycling to school are
  perceived by parents. It may also draw attention to AST as an issue
  affecting children's health which could motivate behaviour change. We
  used natural language processing, including topic modelling, to
  examine how AST is framed in publicly available documents from Ontario
  stakeholders involved in school travel planning. We then compared the
  findings from these documents to a selection of studies on AST and
  explored similarities between the two. We found that AST is framed in
  two ways: i) as a health and environmental issue; and ii) as an
  accessible and feasible transport option for children and parents. The
  frames encourage children and parents to adopt AST given its health
  and environmental benefits by providing resources to support behaviour
  change. The benefits of AST and strategies to support AST that are
  communicated by stakeholders are consistent with the evidence from
  academic research. While these frames present AST in a positive light,
  they may not encourage parents to view current household travel
  behaviours as unhealthy for their children or their community.
  Stakeholders promoting AST in Ontario should further problematize the
  decline of AST and challenge the norm of driving children to school.

  Keywords: active school travel, school travel planning; natural
  language processing; framing analysis; education strategies
  \end{abstract}

@@

\hypertarget{introduction}{%
\section{1. Introduction}\label{introduction}}

Rates of walking and bicycling to school, commonly known as active
school travel (AST), have been declining in Canada and the United States
for decades (Rothman et al., 2018). However, the appetite for AST may be
high in many Canadian communities. For instance, in a study conducted in
Toronto, Ontario, 40\% of children who were driven to school stated they
would like to travel by bicycle instead (Larouche et al., 2016). The
study also reported that less than 3\% of respondents actually cycled,
despite the vast majority having access to a bicycle and living within a
short and bikeable distance to school. Another study based in London,
Ontario, reported similar findings with respect to children's preference
for active travel (Larsen et al., 2012). From a public health
perspective, increasing the use of active modes to school represents a
major opportunity to improve children's health and wellbeing.

The Government of Ontario, in Canada, has striven to promote this trend,
and created a fund in 2017 to support communities across the province to
develop AST initiatives. Green Communities Canada runs the program,
called \emph{Ontario Active School Travel} (G. C. Canada, 2020). As of
December 2021, OAST has awarded over CAD 2 million and provided
resources to over 25 projects across Ontario. Many communities
implemented school travel planning (STP), along with encouraging
activities like walking school buses and developed resources for schools
and parents, among other actions (see Canada, 2021). In this context,
STP is a popular ``school-specific'' intervention led by a facilitator
who brings together a committee of stakeholders from diverse sectors
including education, planning, transportation, and public health to
develop action plans (Buliung et al., 2011; Mammen et al., 2014a). The
five-step process involves identifying barriers to AST based on the
local context and implementing approaches or activities that make AST
safer and more convenient: 1) setup of the program; 2) data collection
and problem identification; 3) action planning; 4) implementation; 5)
monitoring and evaluation (Buliung et al., 2011; Lang et al., 2011).

How AST is framed by STP stakeholders to a target audience, like parents
or the general public, may raise awareness of the issue and influence
how walking and bicycling to school are perceived. Gamson and Modigliani
(1987) define a frame as a ``central organizing idea or story line that
provides meaning'' to a particular phenomenon. A frame can enable
individuals ``to locate, perceive, identify, and label'' information
pertaining to various dimensions of an issue (Goffman, 1974). Therefore,
stakeholders involved in STP efforts can play a role in shaping public
perception about AST in such a way that it attracts greater attention
and is recognized as a ``problem'' that needs to be addressed through
behaviour change or new policies.

After significant investment of human and financial resources to boost
rates of AST in Ontario over the past few years, we ask this central
question: How is AST framed to the public? The following questions are
subsidiary and help us address this main question: What benefits of AST
are communicated to the public? What solutions are proposed to increase
rates of AST?

In this paper, we use natural language processing to examine how three
STP stakeholder groups in Ontario (municipalities, school boards, and
transportation consortia) frame the issue of AST. We assembled a corpus
of texts from public sources that present information for the general
public or parents who are interested in AST. We examine word frequency,
bigrams, and concordances in these selected documents, and also identify
key topics presented by each stakeholder group. We then compare the
findings from these documents to a selection of studies on AST and
explore the extent to which there is concordance between the literature
on AST and materials shared with the public.

\hypertarget{literature-review}{%
\section{2. Literature Review}\label{literature-review}}

\hypertarget{benefits-of-active-school-travel}{%
\subsection{2.1. Benefits of active school
travel}\label{benefits-of-active-school-travel}}

There is value in communicating the benefits of AST to the public in
order to convey the importance of this issue. The desire to increase AST
in Canada is certainly warranted - there is compelling evidence of the
physical and mental health benefits that children who actively commute
to school attain. Faulkner et al. (2009) concluded from their systematic
review that children who travel on foot or by bicycle to school
generally have higher levels of physical activity than their peers who
are driven to school. However, Schoeppe et al. (2015) reported no
association between AST and physical activity in their cross-sectional
study. This relationship could be dose-dependent, meaning that children
would have to travel longer distances to accumulate physical activity. A
walking distance of 1000-1600 metres to school has been found to
contribute to overall levels of physical activity for boys (Faulkner et
al., 2013). The daily routine of travelling to school can be a good
opportunity for children to regularly build physical activity into their
schedule (Mitra, 2013). Research has also shown that using active modes
to school contributes to improve cardiovascular fitness (Børrestad et
al., 2012).

More recently, the literature has explored the link between transport
and children's wellbeing (Waygood et al., 2020), with relevant
applications to the study of travel satisfaction (van den Berg et al.,
2020; Westman et al., 2017b). Being driven to school reduces community
interactions for children, which may negatively impact their social
wellbeing (Waygood and Friman, 2015). In contrast, walking and bicycling
give children opportunities to socialize with friends or siblings
(Michail et al., 2021) and increase positive emotions that contribute to
wellbeing (Ramanathan et al., 2014). This is something that children
seem to value (Zwerts et al., 2010) and that tracks with findings among
university students (Páez and Whalen, 2010). More generally, social
connections through travel appear to be important for wellbeing.
Furthermore, AST can provide opportunities for children to engage with
natural environments (Fusco et al., 2012; Romero, 2015).

\hypertarget{factors-that-influence-active-school-travel-and-mode-choice}{%
\subsection{2.2. Factors that influence active school travel and mode
choice}\label{factors-that-influence-active-school-travel-and-mode-choice}}

Factors that influence AST have been presented and organized using a
socio-ecological model (SEM) (Mitra, 2013) or systems model (Badland et
al., 2016) whereby children's travel behaviour is understood within the
context of household, social, neighbourhood, and policy environments.
The SEM comes from the field of public health and is a useful framework
for understanding complex health behaviours, such as walking and
bicycling habits, This model helps to identify multiple determinants
that need to be addressed by interventions to facilitate behaviour
change. The consensus in the literature is that interventions should
target multiple levels in order to increase levels of healthy and
physically active modes of travel to school (Mitra, 2013). Rather than
describing an extensive list of factors that influence AST and mode
choice in Canada (e.g., Mammen et al., 2012; Mitra, 2013; Rothman et
al., 2018; Wilson et al., 2018), we discuss a few potentially modifiable
determinants that may be targeted for change through STP.

At the individual level, age is often positively associated with AST
(Mammen et al., 2012; Stark et al., 2018; Wilson et al., 2018). There is
evidence that gender is a determinant of AST, and that boys are more
likely to travel using active modes than girls, although this is not a
strong or consistent finding (Rothman et al., 2018; Schoeppe et al.,
2015). Children's mode choice to school is strongly influenced by their
parents' travel behaviours and the complexity of their household's
travel needs (Buliung et al., 2021), as well as daily parental support
(Mah et al., 2017). This indicates that shifting parental perceptions
and habits is important. Convenience and inclement weather have been
cited by parents as barriers to AST (Buliung et al., 2011). Parental
perceptions of the built or school environment (De Meester et al., 2014;
Panter et al., 2010) and their children's skills (Faulkner et al., 2010;
Mammen et al., 2012) also influence mode choice to school.

Distance between home and school is most negatively associated with AST
(Ikeda et al., 2018; Mammen et al., 2012; Pont et al., 2009; Rothman et
al., 2018) with less AST reported among children who have to travel
farther to school - low density suburban development tends to increase
the need for movement, and the reliance on motorized modes (Farber and
Páez, 2011). Many studies have also found that the quality of the built
environment along the route to school and around the school site (Ikeda
et al., 2018; Rothman et al., 2021) and provision of active travel
infrastructure (Chen et al., 2018; Pont et al., 2009) facilitate AST.
Canadian youth report that they feel most safe bicycling on streets in
their neighbourhood or that have low volumes of traffic (T. A. of
Canada, 2020). Finally, concerns about traffic and strangers have been
reported by parents who drive their children to school (Mammen et al.,
2012), which highlights that the volume and speed of cars can be a
concern or deterrent for AST.

\hypertarget{school-travel-planning-in-canada}{%
\subsection{2.3. School travel planning in
Canada}\label{school-travel-planning-in-canada}}

School travel planning (STP) has been implemented in Canada since at
least the late 2000s. Within the STP process, facilitators generally
establish multi-sector committees which intervene at the participating
school through a range of activities related to the 4E's consisting of
\emph{education} strategies, \emph{encouragement} through in-person
events or programs, \emph{engineering} improvements to or around the
school site, and \emph{enforcement} of traffic speeds around schools
(Lang et al., 2011; Mammen et al., 2014b). The first large-scale
evaluation of STP as an intervention in Canada took place at twelve
schools across the country, including four in Ontario, using parental
surveys to measure changes in travel behaviour and perceptions (Buliung
et al., 2011). Following the intervention, active travel rates increased
across all schools and 14\% of parents reported a perceived reduction in
traffic outside the school (Buliung et al., 2011). Assessments of the
efficacy of STP (Buttazzoni et al., 2019; Mammen et al., 2014b) indicate
that it has the potential to encourage behaviour change and adoption of
AST. STP facilitators have recommended that additional time and
resources are needed to improve the efficacy of STP (Mammen et al.,
2015), which highlights that long-term and sustained efforts driven at
the policy level are required to address declining rates of AST.

How AST is framed seems particularly important to shift parental
attitudes and perceptions, given their reported influence on children's
travel mode to school. STP activities heavily focus on education or
encouragement (Buliung et al., 2011; Buttazzoni et al., 2018; Mammen et
al., 2014b), but parents may not always be receptive to the goals of STP
and may be resistant to behaviour change (Buttazzoni et al., 2018).
Parents have been found to express different understandings, language,
and perceptions than planners of how the built environment can influence
school travel (Buliung et al., 2021). This is also true when it comes to
other factors like convenience of different modes to school (Lang et
al., 2011). STP stakeholders must pay special attention to parents'
understanding of the decline of AST as a problem, which may affect their
receptivity to proposed solutions.

The ``central organizing idea or story line'' of AST, to apply the
definition of framing from Gamson and Modigliani (1987), could also
affect broader support in the community. Municipal representatives are
perceived to be instrumental but the involvement of other stakeholder
groups (e.g., busing consortia representatives and local residents) can
be lacking (Buttazzoni et al., 2018). In Ontario, it is reasonable to
say that AST has become a policy issue on the education and public
health agendas, supported, albeit in a modest way, by financial
contributions from the provincial government\footnote{To put the CAD 2
  million invested by the province in the Ontario Active School Travel
  program in perspective, a city such as Hamilton budgeted in excess of
  CAD 91 million on road maintenance and construction in 2016 alone;
  see:
  https://www.hamilton.ca/budget-finance/city-budgets/2016-tax-and-rate-budgets}.
Still, the support from a range of municipal representatives (see
Buttazzoni et al., 2018; Mammen et al., 2015) demonstrates that the AST
issue is on the political agenda. However, it is unknown to what degree
the general public (i.e., local residents) has been exposed to messaging
about AST.

The success of STP interventions would likely depend on parental
judgments of factors that are related to school travel, as well as
support from key policy makers. For example, parents have been found to
view mixed land use as conducive for driving, despite transport planners
viewing neighbourhoods with mixed uses as key for encouraging more
active travel (Buliung et al., 2021). Therefore, stakeholders involved
in STP must make conscious choices about the proposed solutions and
potential benefits of AST that are communicated to parents and the
general public to convey its importance as a policy issue and to
facilitate adoption of AST. Publicly available content about AST needs
to effectively engage multiple audiences on this policy issue, including
parents, children, politicians, and school representatives. This
information should reflect current knowledge from research on school
travel, plus content specific to local factors that influence AST, so
that the challenges of AST are adequately defined and the opportunities
or solutions to address the problems are clear.

\hypertarget{framing-analysis}{%
\subsection{2.4. Framing Analysis}\label{framing-analysis}}

Issues that pertain to public health or wellbeing are often presented to
the public through particular frames to influence perceptions or
behaviours. As previously mentioned, a frame is a ``central organizing
idea or story line that provides meaning'' to a public issue or
phenomenon (Gamson and Modigliani, 1987). Scholars in the field of
political communications have proposed that communicators, such as the
media or institutions, construct the narrative of a frame for policy
positions or public issues in order to activate or restrict a particular
response in the intended audience (Pan and Kosicki, 1993). Organized
groups of stakeholders can employ similar methods to attract attention
to particular issues. Framing can be used to position existing solutions
as suitable to address particular issues (Mah et al., 2014), which may
prevent the public from being aware of other policy approaches that
challenge the status quo. The way policy issues are framed is ultimately
important to understand because it plays a role in either altering or
preserving the existing social perceptions. Finally, the ideas that are
communicated by a particular frame can be either positive or negative
(Waygood and Avineri, 2018). There is evidence that negative framing may
be more effective at motivating individuals to change transport modes to
reduce CO2 emissions (Waygood and Avineri, 2018).

Framing of issues is an important step in developing health policy. An
obvious example over the past decade is the framing of climate change as
a public health issue (e.g., Depoux et al., 2017; Maibach et al., 2010;
Weathers and Kendall, 2016) to increase public engagement and awareness
of the issue. This framing has slowly advanced this issue on public
policy agendas as public attention puts pressure on the policy stream to
adopt frameworks for action. For example, transport planners use
different frames to guide the extent to which transport policies can be
adapted to address climate change. In a recent paper (Reynard et al.,
2021), framing analysis was applied to review the representation of
issues such as mobility and social exclusion in municipal policies from
four western Canadian cities under the current circumstances of climate
change. The authors found four primary frames: ``The Growing City,''
``If You Build It, They Will Come,'' ``Better City for All,'' and a
``the Resilient City'' (Reynard et al., 2021). Each frame presented the
nature, opportunities, and challenges of climate change in different
ways which set the stage for the types of mitigation and adaptation
strategies that cities were proposing to address this issue.

In a similar way, we hypothesize that STP stakeholder groups in Ontario
have framed AST in particular ways to engage multiple audiences on this
policy issue including parents, local residents, school representatives,
and municipal representatives. These groups are likely identifying
points of intervention and potential opportunities to build support for
and encourage AST.

\hypertarget{data}{%
\section{3. Data}\label{data}}

\hypertarget{data-retrieval}{%
\subsection{3.1. Data retrieval}\label{data-retrieval}}

>>>>

\documentclass[preprint, 3p,
authoryear]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%

\usepackage[hyphens]{url}

  \journal{Transportation Research Part A: Policy and
Practice} % Sets Journal name

@@

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[]{natbib}
\bibliographystyle{plainnat}

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={How is active school travel framed in Ontario, Canada?},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)

@@

@@

  \title{How is active school travel framed in Ontario, Canada?}
    \author[McMaster University]{Elise Desjardins%
  %
  }
   \ead{desjae@mcmaster.ca} 
    \author[McMaster University]{Jason Lam%
  %
  }
   \ead{lamj54@mcmaster.ca} 
    \author[University of Alberta]{Darcy Reynard%
  %
  }
   \ead{reynard@ualberta.ca} 
    \author[University of Alberta]{Damian Collins%
  %
  }
   \ead{damian.collins@ualberta.ca} 
    \author[Polytechnique Montreal]{E. Owen D. Waygood%
  %
  }
   \ead{owen.waygood@polymtl.ca} 
    \author[McMaster University]{Antonio Paez%
  \corref{cor1}%
  \fnref{1}}
   \ead{paezha@mcmaster.ca} 
      \affiliation[McMaster University]{School of Earth, Environment and
Society, 1280 Main St.~W, Hamilton, Ontario, Canada L8S 4K1}
    \affiliation[University of Alberta]{Department of Earth and
Atmospheric Sciences, 1-26 Earth Sciences Building, Edmonton, Alberta,
Canada T6G 2E3}
    \affiliation[Polytechnique Montreal]{Department of Civil, Geological
and Mining Engineering, 2500, chemin de Polytechnique Montreal,
Montreal, Quebec, Canada H3T 1J4}
    \cortext[cor1]{Corresponding author}
    \fntext[1]{Corresponding author.}

  \begin{abstract}
  Active school travel (AST) has been increasingly encouraged by various
  stakeholders in Ontario, Canada through efforts such as school travel
  planning. Education strategies like workshops or resources that
  promote AST are commonly implemented. The framing of AST through such
  strategies may influence how walking and bicycling to school are
  perceived by parents. It may also draw attention to AST as an issue
  affecting children's health which could motivate behaviour change. We
  used natural language processing, including topic modelling, to
  examine how AST is framed in publicly available documents from Ontario
  stakeholders involved in school travel planning. We then compared the
  findings from these documents to a selection of studies on AST and
  explored similarities between the two. We found that AST is framed in
  two ways: i) as a health and environmental issue; and ii) as an
  accessible and feasible transport option for children and parents. The
  frames encourage children and parents to adopt AST given its health
  and environmental benefits by providing resources to support behaviour
  change. The benefits of AST and strategies to support AST that are
  communicated by stakeholders are consistent with the evidence from
  academic research. While these frames present AST in a positive light,
  they may not encourage parents to view current household travel
  behaviours as unhealthy for their children or their community.
  Stakeholders promoting AST in Ontario should further problematize the
  decline of AST and challenge the norm of driving children to school.
  \end{abstract}
    \begin{keyword}
    active school travel \sep school travel planning \sep natural
language processing \sep 
    framing analysis
  \end{keyword}

@@

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Walking and bicycling to school, commonly known as active school travel
(AST), have declined in North America for decades
\citep{rothmanDeclineActiveSchool2018}, but the appetite for these modes
is beginning to rise in many Canadian communities. For instance,
research in Toronto, Ontario found that 40\% of children who were driven
to school would like to travel by bicycle instead
\citep{laroucheRatherBikeSchool2016}. Another study in London, Ontario,
reported similar findings with respect to children's preference for
active travel \citep{larsenRouteBasedAnalysisCapture2012}. From a public
health perspective, increasing active travel to school represents a
major opportunity to improve children's health and wellbeing.

To promote this trend, in 2017 the Government of Ontario began to
support communities develop AST initiatives through a program called
\emph{Ontario Active School Travel} (OAST). By 2021, the program, run by
Green Communities Canada had awarded over CAD 2 million and provided
resources to over 25 projects across Ontario
\citep{GreenCommunities2016}. A popular intervention in many communities
is school travel planning (STP) \citep[see][]{GreenCommunitiesProjects}.
In this context, STPs are ``school-specific'' interventions led by a
facilitator who works with a committee of stakeholders from diverse
sectors including education, planning, transportation, and public health
to develop action plans
\citep{buliungSchoolTravelPlanning2011, mammenSchoolTravelPlanning2014}.
The five-step process involves identifying barriers to AST and
implementing solutions or activities to make AST safer and more
convenient: 1) setup of the program; 2) data collection and problem
identification; 3) action planning; 4) implementation; 5) monitoring and
evaluation
\citep{buliungSchoolTravelPlanning2011, langUnderstandingModalChoice2011}.

Framing AST by STP stakeholders to a target audience (e.g., parents or
the general public) can raise awareness of the relevant issues and
influence how active travel to school is perceived. Gamson and
Modigliani \citeyearpar{gamsonChangingCulture1987} define a frame as a
``central organizing idea or story line that provides meaning'' to a
particular phenomenon. A frame can enable individuals ``to locate,
perceive, identify, and label'' information pertaining to various
dimensions of an issue \citep{goffmanFrameAnalysisEssay1974}. Therefore,
stakeholders involved in STP efforts can play a role in shaping public
perception about AST to attract greater attention and recognition as a
``problem'' that needs to be addressed through behaviour change or new
policies.

After significant investment of human and financial resources to boost
rates of AST in Ontario over the past few years, in this paper we ask
this central question: How is AST framed to the public in Ontario? The
following questions are subsidiary and help us address this main
question: What benefits of AST are communicated to the public? What
solutions are proposed to increase AST?

We approach this question by means of natural language processing, and
examine how three STP stakeholder groups in Ontario (municipalities,
school boards, and transportation consortia) frame AST. We assemble a
corpus of texts from public sources that present information for the
general public or parents who are interested in AST. We examine word
frequency, bigrams, and concordances in these selected documents, and
also identify key topics presented by each stakeholder group. We then
compare the findings from these documents to a selection of studies on
AST and explore the extent to which there is concordance between the
scholarly literature on AST and materials shared with the public.

\hypertarget{literature-review-ast}{%
\section{Literature Review: AST}\label{literature-review-ast}}

The desire to increase AST in Canada is warranted given its potential
physical and mental health benefits. Faulkner et al.
\citeyearpar{faulknerActiveSchoolTransport2009} concluded from their
systematic review that children who travel on foot or by bicycle to
school generally have higher levels of physical activity than their
peers who are driven to school. This relationship could be
dose-dependent, and contingent on sufficiently long trips to school to
accumulate physical activity. A walking distance of 1000-1600 metres to
school has been found to contribute to overall levels of physical
activity for boys \citep{faulknerSchoolTravelChildren2013}. The daily
routine of travelling to school can be a good opportunity for children
to regularly build physical activity into their schedule
\citep{mitraIndependentMobilityMode2013}.

More recently, the link between transport and children's wellbeing has
attracted interest too \citep{waygoodTransportChildrenWellbeing2020}.
Being driven to school reduces community interactions for children,
which may negatively impact their social wellbeing
\citep{waygoodChildrenTravelIncidental2015}. In contrast, walking and
bicycling give children opportunities to socialize
\citep{michailChildrenExperiencesTheir2021} and increase positive
emotions that contribute to wellbeing
\citep{ramanathanHappinessMotionEmotions2014a}, a feeling that children
seem to value \citep{zwertsHowChildrenView2010}. AST can also provide
opportunities for children to engage with natural environments
\citep{fuscoUnderstandingChildrenPerceptions2012, romeroChildrenExperiencesEnjoyment2015}.

Factors that influence AST have been presented and organized using a
socio-ecological model (SEM) \citep{mitraIndependentMobilityMode2013} or
systems model \citep{badlandDevelopmentSystemsModel2016} whereby
children's travel behaviour is understood within the context of
household, social, neighbourhood, and policy environments. The SEM helps
to identify multiple determinants that need to be addressed by
school-based interventions to increase AST, with the consensus being
that interventions should target multiple levels
\citep{mitraIndependentMobilityMode2013}. Rather than describing an
extensive list of factors that influence AST and mode choice in Canada
\citep[e.g.,][]{mammenUnderstandingDriveEscort2012, mitraIndependentMobilityMode2013, rothmanDeclineActiveSchool2018, wilsonUnderstandingChildParent2018},
we discuss a few potentially modifiable determinants that may be
targeted for change through STP.

Children's mode choice to school is strongly influenced by their
parents' travel behaviours and the complexity of their household's
travel needs \citep{buliungLivingJourneySchool2021}, as well as daily
parental support \citep{mahDoesParentalSupport2017a}. This indicates
that shifting parental perceptions and habits is important. Convenience
and inclement weather have been cited by parents as barriers to AST
\citep{buliungSchoolTravelPlanning2011}. Parental perceptions of the
built or school environment
\citep{demeesterParentalPerceivedNeighborhood2014, panterAttitudesSocialSupport2010}
and their children's skills
\citep{mammenUnderstandingDriveEscort2012, faulknerWhatQuickestEasiest2010}
also influence mode choice to school.

Distance between home and school is most negatively associated with AST
\citep{ikedaAssociationsChildrenActive2018, mammenUnderstandingDriveEscort2012, pontEnvironmentalCorrelatesChildren2009, rothmanDeclineActiveSchool2018}
with less AST reported among children who have to travel farther to
school. Many studies have also found that the quality of the built
environment along the route to school and around the school site
\citep{ikedaAssociationsChildrenActive2018, rothmanActiveSchoolTransportation2021}
and provision of active travel infrastructure
\citep{chenPromotingActiveStudent2018, pontEnvironmentalCorrelatesChildren2009}
facilitate AST. Canadian youth report that they feel most safe bicycling
on streets in their neighbourhood or that have low volumes of traffic
\citep{TACbikeinfra2020}. Finally, concerns about traffic and strangers
have been reported by parents who drive their children to school
\citep{mammenUnderstandingDriveEscort2012}, which highlights that the
volume and speed of cars can be a concern or deterrent for AST.

\hypertarget{school-travel-planning-in-canada-and-relevance-of-framing-analysis}{%
\section{School travel planning in Canada and relevance of framing
analysis}\label{school-travel-planning-in-canada-and-relevance-of-framing-analysis}}

School travel planning (STP) has been implemented in Canada since at
least the late 2000s. Within the STP process, facilitators establish
multi-sector committees which intervene at the participating school
through a range of activities consisting of \emph{{[}E{]}ducation}
strategies, \emph{{[}E{]}ncouragement} through in-person events or
programs, \emph{{[}E{]}ngineering} improvements to or around the school
site, and \emph{{[}E{]}nforcement} of traffic speeds around schools
\citep[or
4Es,][]{langUnderstandingModalChoice2011, mammenActiveSchoolTravel2014}.
Following the first large-scale evaluation of STP as an intervention in
Canada, AST rates increased and 14\% of parents reported a perceived
reduction in traffic outside the school
\citep{buliungSchoolTravelPlanning2011}. Assessments of the efficacy of
STP
\citep{buttazzoniPromotingActiveSchool2019, mammenActiveSchoolTravel2014}
indicate its potential to encourage behaviour change and adoption of
AST.

By providing a ``central organizing idea or story line''
\citep{gamsonChangingCulture1987} framing of AST seems particularly
important to shift parental attitudes and perceptions, given their
reported influence on children's travel mode to school. Scholars in the
field of political communications have proposed that communicators, such
as the media or institutions, construct the narrative of a frame for
policy positions or public issues in order to activate or restrict a
particular response in the intended audience
\citep{panFramingAnalysisApproach1993}. Organized groups of stakeholders
can employ similar methods to attract attention to particular issues.
Framing can be used to position existing solutions as suitable to
address particular issues \citep{mahFramecriticalPolicyAnalysis2014},
which may prevent the public from being aware of other policy approaches
that challenge the status quo. The way policy issues are framed is
ultimately important to understand because it plays a role in either
altering or preserving the existing social perceptions. Finally, the
ideas that are communicated by a particular frame can be either positive
or negative \citep{waygoodCO2ValenceFraming2018}. There is evidence that
negative framing may be more effective at motivating individuals, for
example, to change transport modes to reduce CO2 emissions
\citep{waygoodCO2ValenceFraming2018}.

Framing of AST could also affect broader support in the community.
Municipal representatives are perceived to be instrumental but the
involvement of other stakeholder groups (e.g., busing consortia
representatives and local residents) can be lacking
\citep{buttazzoniSupportingActiveSchool2018}. In Ontario, it is
reasonable to say that AST has become a policy issue on the education
and public health agendas, supported, albeit in a modest way, by
financial contributions from the provincial government. Still, the
support from a range of municipal representatives
\citep[see][]{buttazzoniSupportingActiveSchool2018, mammenPuttingSchoolTravel2015}
demonstrates that the AST issue is on the political agenda.

Since STP activities heavily focus on education or encouragement
\citep{buliungSchoolTravelPlanning2011, mammenActiveSchoolTravel2014, buttazzoniSupportingActiveSchool2018},
it is important that parents and communities are receptive to the goals
of STP \citep{buttazzoniSupportingActiveSchool2018}. Parents have been
found to express different understandings, language, and perceptions
than planners of how the built environment can influence school travel
\citep{buliungLivingJourneySchool2021}. This is also true when it comes
to other factors like convenience of different modes to school
\citep{langUnderstandingModalChoice2011}. STP stakeholders must pay
special attention to parents' understanding of the decline of AST as a
problem, which may affect their receptivity to proposed solutions.

The success of STP interventions would likely depend on parental
judgments of factors that are related to school travel, as well as
support from key policy makers. For example, parents have been found to
view mixed land use as conducive for driving, despite transport planners
viewing neighbourhoods with mixed uses as key for encouraging more
active travel \citep{buliungLivingJourneySchool2021}. Therefore,
stakeholders involved in STP must make conscious choices about the
proposed solutions and potential benefits of AST that are communicated
to parents and the general public through publicly available content to
convey its importance as a policy issue and to facilitate adoption of
AST.

\hypertarget{data}{%
\section{Data}\label{data}}

\hypertarget{data-retrieval}{%
\subsection{Data retrieval}\label{data-retrieval}}

\hypertarget{policy-documents}{%
\subsubsection{Policy documents}\label{policy-documents}}

We assembled a collection of publicly available documents that were
sourced online from the main stakeholder groups involved in STP
initiatives in Ontario: i) school boards (public or Catholic and
English-speaking only); ii) municipal governments; and iii)
transportation consortia. The latter involve a collaboration between
municipal regions and school boards to deliver more efficient and timely
transportation services to schools in each region. Non-profit
organizations, police services, and advocacy groups are other
stakeholders who often play a role in supporting AST and/or STP, but
this study does not include any documents from these groups because they
do not consistently participate in STP initiatives in Ontario.

The search was guided first by a list of all English public and Catholic
school boards across Ontario. The websites of each school board were
manually searched for pages related to school transport or travel. Any
pages relevant to these topics were downloaded. Next, we searched
municipal government and transportation consortia websites. These were
identified based on geographic area. Likewise, webpages related to
active transport or school travel were downloaded.

Webpages from STP stakeholder groups were included in our analysis if
they were findable. This primary criterion was important since our
analysis pertains to how such issues are framed to the general public.
Thus, we included only webpages that were readily accessible, which we
defined as requiring no more than four links from the initial Google
search to reach.

The initial corpus of documents from STP stakeholder groups included 69
relevant webpages. We refer to these as policy+practice documents
throughout the paper. It is important to note that school boards,
municipalities, and transportation consortia may or may not publish
information about their involvement in AST and STP efforts on their
respective websites or in policy documents. Search results are
summarized in Table \ref{tab:policy-documents}.
